---
title: "MM916 Regression Projet"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2023-11-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(GGally)
# Read in the data
real_est <- read.csv("real_est.csv", header=TRUE)
```


```{r Summary Statistics}
summary(real_est)
var(real_est$Price)
```

Finding the relationship between numeric predictor variables and Price

```{r Scatterplot for Price vs other numeric variables}
par(mfrow=c(3, 3))
par(mar=c(4, 4, 2, 2))
variables <- c('Bedroom', 'Space', 'Room', 'Lot', 'Tax', 'Bathroom', 'Garage')
for (var in variables) {
 plot(real_est[[var]], real_est$Price,
 ylab = 'Price',
 xlab = var,
 col = 'darkgreen',
 pch = 19)
 lines(lowess(real_est[[var]], real_est$Price), col = 'red')
}
ggsave("plot.png", width = 6, height = 4)

```


Finding the relationship between categorical variable with other variables and target variable

```{r Density plot for Condition interaction}
library(ggplot2)
library(patchwork)

options(repr.plot.width = 14, repr.plot.height = 12, repr.plot.res = 100)
options(repr.plot.width = 20, repr.plot.height = 12, repr.plot.res = 100)

numeric_vars <- c("Price", "Room", "Bedroom", "Space", "Garage", "Lot", "Tax", "Bathroom")

density_plots <- lapply(numeric_vars, function(var) {
  ggplot(real_est, aes(x = !!as.name(var), fill = factor(Condition))) +
    geom_density(alpha = 0.5) +
    labs(title = paste( var), x = var, y = "Density") +
    scale_fill_discrete(name = "Condition")
})

plots_arranged <- wrap_plots(density_plots, nrow = 3) +
  plot_layout(guides = "collect", widths = c(6, 6, 4))
plots_arranged

ggsave("condition.png", width = 6, height = 4)

```


The below correlation matrix is used to analyze the relationship between target variable Price and other independent variables

```{r Correlation analysis}
library(corrplot)

# Calculate the correlation matrix
correlation_matrix <- cor(real_est)

# Define the order of variables
desired_order <- c("Price", "Bedroom", "Space", "Room", "Lot", "Tax", "Garage", "Bathroom", "Condition")

# Reorder the correlation matrix based on desired order
reordered <- correlation_matrix[desired_order, desired_order]

# Use a divergent color scale for the correlation plot
corrplot(
  reordered,
  method = "color",
  type = "upper",
  col = colorRampPalette(c("darkblue", "white"))(100),  # Divergent color scale
  tl.col = "black",
  tl.srt = 45,
  addCoef.col = "black",
  number.cex = 0.7,
  order = "original",
  addrect = 2,
  diag = FALSE,
  mar = c(0, 0, 2, 0)
)

```

A model is build using the above correlation matrix.
```{r Initial Linear Model}
real_est_mod <- lm(Price~ Bedroom+Space+Room+Lot+Tax+Bathroom+Garage+Condition, data=real_est)
summary(real_est_mod)
```




```{r Checking model assumptions}
par(mfrow=c(2,2), mar=c(4.5,4,2,2))
plot(real_est_mod)
require(ggfortify)
autoplot(real_est_mod)

```


```{r Independence of error - Residuals vs Fitted Plot}

par(mfrow = c(3, 2), mar = c(4, 4, 2, 1))  # Increase bottom margin to 2

# Plotting studentized residuals against each independent variable
for (i in names(real_est)[2:7]) {
  plot(real_est[[i]], rstudent(real_est_mod),
       xlab = i, ylab = "Studentized Residuals",
       main = paste("Residuals vs", i),
       col = "coral")
}
```

```{r BoxCox transformation}
library(MASS)
bc_real_est <- boxcox(real_est_mod)
bc_bedroom <- boxcox(lm(Bedroom ~ 1, data = real_est))
bc_room <- boxcox(lm(Room ~ 1, data = real_est))
bc_space <- boxcox(lm(Space ~ 1, data = real_est))
bc_lot <- boxcox(lm(Lot ~ 1, data = real_est))
bc_tax <- boxcox(lm(Tax ~ 1, data = real_est))
bc_bathroom <- boxcox(lm(Bathroom ~ 1, data = real_est))
bc_garage <- boxcox(lm((Garage+0.01) ~ 1, data = real_est))

optimal_lambda_bedroom <- bc_bedroom$x[which.max(bc_bedroom$y)]
optimal_lambda_room <- bc_room$x[which.max(bc_room$y)]
optimal_lambda_space <- bc_space$x[which.max(bc_space$y)]
optimal_lambda_lot <- bc_lot$x[which.max(bc_lot$y)]
optimal_lambda_tax <- bc_tax$x[which.max(bc_tax$y)]
optimal_lambda_bathroom <- bc_bathroom$x[which.max(bc_bathroom$y)]
optimal_lambda_garage <- bc_garage$x[which.max(bc_garage$y)]

# Display optimal lambda values
optimal_lambda_values <- data.frame(
  Variable = c("Bedroom", "Room", "Space", "Lot", "Tax", "Bathroom", "Garage"),
  Optimal_Lambda = c(
    optimal_lambda_bedroom,
    optimal_lambda_room,
    optimal_lambda_space,
    optimal_lambda_lot,
    optimal_lambda_tax,
    optimal_lambda_bathroom,
    optimal_lambda_garage
  )
)

optimal_lambda_values$Optimal_Lambda <- round(optimal_lambda_values$Optimal_Lambda,1)
optimal_lambda_values

```




```{r Transforming variables}
#transforming the variables according the lambda values
transformed_real_est <- real_est

transformed_real_est$Bedroom <- log(real_est$Bedroom)
transformed_real_est$Room <- real_est$Room^(-1)
transformed_real_est$Space <- real_est$Space^(-1)
transformed_real_est$Lot <- real_est$Lot^(-2)
transformed_real_est$Tax <- real_est$Tax^(-1)
transformed_real_est$Bathroom <- real_est$Bathroom^(-1)
transformed_real_est$Garage <- log((real_est$Garage+0.01))

head(transformed_real_est)
```

```{r Distribution of variables after transformation}
par(mfrow=c(2, 2))
variables <- c('Price', 'Space', 'Garage', 'Lot', 'Bedroom', 'Bathroom', 'Room', 'Tax')
for (var in variables) {
  hist(transformed_real_est[[var]], main="", xlab=var, ylab="Frequency", col="skyblue", border="black", breaks=20)
}
```

```{r }
transformed_real_est_mod <- lm(Price~ Bedroom+Space+Room+Lot+Tax+Bathroom+Garage+Condition, data=transformed_real_est)
summary(transformed_real_est_mod)
```

```{r transformation without changing Garage}
transformed_real_est1 <- real_est

transformed_real_est1$Bedroom <- log(real_est$Bedroom)
transformed_real_est1$Room <- real_est$Room^(-1)
transformed_real_est1$Space <- real_est$Space^(-1)
transformed_real_est1$Lot <- real_est$Lot^(-2)
transformed_real_est1$Tax <- real_est$Tax^(-1)
transformed_real_est1$Garage <- real_est$Garage
transformed_real_est1$Bathroom <- real_est$Bathroom
```

```{r transformed model considered henceforth}
transformed_real_est1_mod <- lm(Price~ Bedroom+Space+Room+Lot+Tax+Bathroom+Garage+Condition, data=transformed_real_est)
summary(transformed_real_est1_mod)
```


```{r Checking model assumptions after transformation}
par(mfrow=c(2,2), mar=c(4.5,4,2,2))
plot(transformed_real_est1_mod)
require(ggfortify)
autoplot(transformed_real_est1_mod)

```



```{r}
library(car)
vif_results <- car::vif(transformed_real_est1_mod)
vif_results
```

```{r Handling categorical variables}
#categorical interaction:
categorical_full_mod <- lm(Price ~ Bedroom + Space + Room + Lot + Tax + Bathroom + Garage + Condition + Bedroom:Condition + 
                           Space:Condition + Room:Condition +  Lot:Condition + Tax:Condition + Bathroom:Condition +Garage:Condition, data = transformed_real_est1)
summary(categorical_full_mod)
```

```{r Handling categorical variables removing insignifcant variables}
#categorical interaction:
categorical_full_mod2 <- lm(Price ~ Bedroom + Space + Room + Lot + Tax + Bathroom + Garage + Condition +  Bathroom:Condition, data = transformed_real_est1)
summary(categorical_full_mod2)
AIC(categorical_full_mod2)
```
```{r}
anova(categorical_full_mod, categorical_full_mod2)
```


```{r Backward Elimination}
full_mod <- lm(Price ~ Bedroom + Space + Room + Lot + Tax + Bathroom + Garage + Condition +  Bathroom:Condition, data = transformed_real_est1)
drop1(full_mod, scope = ~Bedroom + Space + Room + Lot + Tax + Bathroom + Garage + Condition +  Bathroom:Condition, data = transformed_real_est1, test = "F")
```

```{r}
#removing insignificant Room variable
back_mod1 <- lm(Price ~ Bedroom + Space  + Lot + Tax + Bathroom + Garage + Condition +  Bathroom:Condition, data = transformed_real_est1)
drop1(back_mod1, scope = ~Bedroom + Space  + Lot + Tax + Bathroom + Garage + Condition +  Bathroom:Condition, data = transformed_real_est1, test = "F")
```
```{r}
summary(back_mod1)
```
Stopping the variable selection as all variables are significant

```{r checking model score}
AIC(full_mod)
AIC(back_mod1)
```

```{r Step Method}
library(car)
step_mod <- step(categorical_full_mod2)
summary(step_mod)
car::vif(step_mod, type = 'predictor') #GVIF^(1/(2*DF)) should be less than 2
```

```{r}
summary(step_mod)
```

```{r}
AIC(real_est_mod)
AIC(transformed_real_est1_mod)
AIC(categorical_full_mod2)
AIC(back_mod1)
AIC(step_mod)
```
```{r}
summary(real_est_mod)
summary(transformed_real_est1_mod)
summary(categorical_full_mod2)
summary(back_mod1)
summary(step_mod)
```
```{r}
final_mod <- step_mod
summary(final_mod)
```
```{r}
par(mfrow=c(2,2), mar=c(4.5,4,2,2))
plot(final_mod)
require(ggfortify)
autoplot(final_mod)
```

```{r Multicollinearty}
vif_results <- car::vif(final_mod)

# View the VIF values
vif_results

```

```{r}
final_model <- lm(Price ~ Bedroom + Space + Lot + Tax + Bathroom + Garage + Condition + Bathroom:Condition, data = transformed_real_est1)

coefficients <- coef(final_model)
formula <- as.formula(final_model)

# Getting variable names
variables <- all.vars(formula)[-1]  # Excluding the intercept

# Constructing the equation with variable names
equation <- paste("Price =", paste(coefficients[2:length(coefficients)], variables, sep = " * ", collapse = " + "), collapse = " ")
intercept <- coefficients[1]  # Intercept value
equation_with_intercept <- paste(equation, intercept, sep = " + ")
equation_with_intercept


```
